---
title: "Banking Dataset — Data Mining Analysis (Group 1)"
author: "Group 1"
date: "`r format(Sys.Date(), '%Y-%m-%d')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(tidyverse)
library(lubridate)
```

# Problem Definition

- Dataset: banking_realistic.csv (Group 1)
- Problem: Predict `LoanStatus` (Approved / Rejected) and analyze factors affecting loan decisions.
- Objective: Build a classification model for `LoanStatus`, provide feature importance, and perform exploratory analysis on Income, Expenditure, Score, Age, Category, and Review.

# Data Understanding

```{r read-data}
df <- read_csv(file.path('Project_datasets','banking_realistic.csv'))
glimpse(df)
cat('Rows:', nrow(df), 'Cols:', ncol(df), '\n')
head(df)
```

## Basic summary and missing values

```{r summary}
summary(df)
sapply(df, function(x) sum(is.na(x)))
```

## Convert types

```{r types}
df <- df %>%
  mutate(Date = ymd(Date),
         LoanStatus = as.factor(LoanStatus),
         Category = as.factor(Category),
         Review = as.factor(Review))

glimpse(df)
```

## First plots (distributions)

```{r plots}
ggplot(df, aes(x = Income)) + geom_histogram(bins=30) + theme_minimal()
ggplot(df, aes(x = Expenditure)) + geom_histogram(bins=30) + theme_minimal()
ggplot(df, aes(x = Score)) + geom_histogram(bins=20) + theme_minimal()
```

<!-- Next steps: Data cleaning, feature engineering, modeling (classification), clustering, text analysis on `Review`. -->
